\textbf{SPECFEM3D}\\ \textbf{User Manual}

### 2012-09-28 doc/USER_MANUAL/manual_SPECFEM3D.tex
adds a check on the use of the ABSORB_INSTEAD_OF_FREE_SURFACE flag
adds a set of FORCESOLUTION files to set force source parameters, used instead of CMTSOLUTION files if "USE_FORCE_POINT_SOURCE = .true".

\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a force source

@@

its norm is ignored and the norm of the force used is \texttt{FACTOR\_FORCE\_SOURCE} times the source time function.

@@

The solver needs three input files in the \texttt{in\_data\_files/} directory
to run:
\begin{description}
\item [{\texttt{Par\_file}}] the main parameter file which was discussed in detail in the previous
Chapter~\ref{cha:Creating-Distributed-Databases},
\item [{\texttt{CMTSOLUTION}}] the earthquake source parameter file, and
\item [{\texttt{STATIONS}}] the stations file.
\end{description}

@@

\begin{itemize}
\item the simulation type control parameters: \texttt{SIMULATION\_TYPE}
and \texttt{SAVE\_FORWARD}
\item the time step parameters \texttt{NSTEP} and \texttt{DT}
\item the absorbing boundary control parameter \texttt{ABSORBING\_CONDITIONS}
\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
\item the ShakeMap\textregistered option \texttt{CREATE\_SHAKEMAP}
\item the output information parameters \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
and \texttt{NTSTEP\_BETWEEN\_OUTPUT\_}~\\
\texttt{SEISMOS}
\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
\end{itemize}
Any other change to the \texttt{Par\_file} implies rerunning both
the database generator \texttt{xgenerate\_databases} and the solver \texttt{xspecfem3D}.

@@

@@

Having set the \texttt{Par\_file} parameters, and having provided
the \texttt{CMTSOLUTION} and \texttt{STATIONS} files, you are now
ready to launch the solver! This is most easily accomplished based
upon the \texttt{go\_solver} script (See Chapter \ref{cha:Scheduler}
for information about running through a scheduler, e.g., LSF). You
may need to edit the last command at the end of the script that invokes
the \texttt{mpirun} command. The \texttt{runall} script compiles and
runs both \texttt{xgenerate\_databases} and \texttt{xspecfem3D} in sequence. This is a safe approach that
ensures using the correct combination of distributed database output and solver
input.

@@

For the same model, one can rerun the solver for different events
by simply changing the \texttt{CMTSOLUTION} file, or for different
stations by changing the \texttt{STATIONS} file. There is no need
to rerun the \texttt{xgenerate\_databases} executable. Of course it is best to include as many stations
as possible, since this does not add to the cost of the simulation.

@@

The SPECFEM3D software code internally uses Cartesian coordinates.
The given locations for sources and receiver locations thus may get converted.
Sources and receiver locations are read in from the
\texttt{CMTSOLUTION} and \texttt{STATIONS} file.
Note that e.g. the \texttt{CMTSOLUTION} file denotes the location by "longitude/latitude/depth". We read in
longitude as $x$ coordinate, latitude as $y$ coordinate.

@@

The value for depth (given in $km$ in \texttt{CMTSOLUTION}) or burial depth (given in $m$ in \texttt{STATIONS})
is evaluated with respect to the surface of the mesh at the specified $x$/$y$ location to find the corresponding $z$ coordinate.
It is possible to use this depth value directly as $z$ coordinate by
changing the flag \texttt{USE\_SOURCES\_RECVS\_Z} to \texttt{.true.} in the file \texttt{constants.h} located in the \texttt{src/shared/} subdirectory.

>>>>

\item [{\texttt{USE\_FORCE\_POINT\_SOURCE}}] Turn this flag on to use a \texttt{FORCESOLUTION} force source

@@

its norm is ignored and the norm of the force used is the factor force source times the source time function.

@@

The solver needs three input files in the \texttt{in\_data\_files/} directory
to run:
\begin{description}
\item [{\texttt{Par\_file}}] the main parameter file which was discussed in detail in the previous
Chapter~\ref{cha:Creating-Distributed-Databases},
\item [{\texttt{CMTSOLUTION}} or {\texttt{FORCESOLUTION}}] the earthquake source parameter file or the force source parameter file or, and
\item [{\texttt{STATIONS}}] the stations file.
\end{description}

@@

\begin{itemize}
\item the simulation type control parameters: \texttt{SIMULATION\_TYPE}
and \texttt{SAVE\_FORWARD}
\item the time step parameters \texttt{NSTEP} and \texttt{DT}
\item the absorbing boundary control parameter \texttt{ABSORBING\_CONDITIONS} on condition that the\\ 
  \texttt{ABSORB\_INSTEAD\_OF\_FREE\_SURFACE} flag remains unmodified after running the databases generation. 
\item the movie control parameters \texttt{MOVIE\_SURFACE}, \texttt{MOVIE\_VOLUME},
and \texttt{NTSTEPS\_BETWEEN\_FRAMES}
\item the ShakeMap\textregistered option \texttt{CREATE\_SHAKEMAP}
\item the output information parameters \texttt{NTSTEP\_BETWEEN\_OUTPUT\_INFO}
and \texttt{NTSTEP\_BETWEEN\_OUTPUT\_}~\\
\texttt{SEISMOS}
\item the \texttt{PRINT\_SOURCE\_TIME\_FUNCTION} flags
\end{itemize}
Any other change to the \texttt{Par\_file} implies rerunning both
the database generator \texttt{xgenerate\_databases} and the solver \texttt{xspecfem3D}.

@@

\vspace{1cm}

\noindent The \texttt{FORCESOLUTION} should be edited in the following way:

\begin{itemize}
\item Set the \texttt{time shift} parameter equal to $0.0$ (the solver
will not run otherwise.) The time shift parameter would simply apply
an overall time shift to the synthetics, something that can be done
in the post-processing (see Section \ref{sec:Process-data-and-syn}).
\item Set the latitude, longitude, depth of the source in geographical
  coordinates.
\item Set the magnitude of the force source.
\item Set the components of a (non-unitary) direction vector for the
  force source in the East/North/Vertical basis (see Appendix A for
  the orientation of the reference frame).
\end{itemize}

@@

Having set the \texttt{Par\_file} parameters, and having provided the
\texttt{CMTSOLUTION} (or the \texttt{FORCESOLUTION}) and
\texttt{STATIONS} files, you are now ready to launch the solver! This
is most easily accomplished based upon the \texttt{go\_solver} script
(See Chapter \ref{cha:Scheduler} for information about running through
a scheduler, e.g., LSF). You may need to edit the last command at the
end of the script that invokes the \texttt{mpirun} command. The
\texttt{runall} script compiles and runs both
\texttt{xgenerate\_databases} and \texttt{xspecfem3D} in
sequence. This is a safe approach that ensures using the correct
combination of distributed database output and solver input.

@@

For the same model, one can rerun the solver for different events by
simply changing the \texttt{CMTSOLUTION} or \texttt{FORCESOLUTION}
file, or for different stations by changing the \texttt{STATIONS}
file. There is no need to rerun the \texttt{xgenerate\_databases}
executable. Of course it is best to include as many stations as
possible, since this does not add to the cost of the simulation.

@@

The SPECFEM3D software code internally uses Cartesian coordinates.
The given locations for sources and receiver locations thus may get
converted.  Sources and receiver locations are read in from the
\texttt{CMTSOLUTION} (or \texttt{FORCESOLUTION}) and \texttt{STATIONS}
files.  Note that e.g. the \texttt{CMTSOLUTION} and
\texttt{FORCESOLUTION} files denotes the location by
"longitude/latitude/depth". We read in longitude as $x$ coordinate,
latitude as $y$ coordinate.

@@

The value for depth (given in $km$ in \texttt{CMTSOLUTION} and
\texttt{FORCESOLUTION}) or burial depth (given in $m$ in
\texttt{STATIONS}) is evaluated with respect to the surface of the
mesh at the specified $x$/$y$ location to find the corresponding $z$
coordinate.  It is possible to use this depth value directly as $z$
coordinate by changing the flag \texttt{USE\_SOURCES\_RECVS\_Z} to
\texttt{.true.} in the file \texttt{constants.h} located in the
\texttt{src/shared/} subdirectory.
